Widely Bandwidth-Tunable Broadband Optical Filter on Silicon

### 2015-05-19 CacadedTunableFilter.tex
Temp Inset and Fixes bib

A broadband filter design for a coarse WDM on the 220-nm SOI platform \cite{shi2013siliconCWDM} is adopted for individual contra-DCs.  
In each contra-DC, the widths of the waveguides are of 560 and 440 nm.
This high asymmetry ensures a negligible forward coupling. 
The gratings are formed by sidewall corrugations in strip waveguides with a pitch of 312 nm. 
The spacing between the waveguides varies between 65 and 135 nm with a Gaussian profile as shown in Fig.~\ref{fig:SEM} for efficient sidelobe suppression, for which details can be found in \cite{shi2013siliconCWDM}.
The spectral responses of each contra-DC is calculated using coupled mode theory, following the same procedure as in  \cite{shi2013siliconContraDC}, with two additions: apodization and noise simulation.
%The first part in predicting the contra-directional coupler spectrum is to compute the dielectric variation caused by the perturbation $\Delta\varepsilon(x,y)$ which is $\varepsilon_0/\pi*(n_{max}(x,y)^2-n_{min}(x,y)^2)$ where the $1/\pi$ factor is the first Fourier expansion coefficient in the case of a square shaped perturbation. 
%One also needs the mode distribution in each waveguide  $E_1(x,y)$ and $E_2(x,y)$ at the average point of the perturbation, which can be obtained using a 2D eigenmode solver such as Lumerical MODE. 
%The couplings $\kappa_{11}$, $\kappa_{12}$ and $\kappa_{22}$  can then be calculated from
%\begin{equation}\label{eq:coupling}
%\kappa_{nm}=\frac{\omega}{4}\iint \textbf{E}^*_n(x,y)\Delta\varepsilon_1(x,y) \textbf{E}_m(x,y) \diff x \diff y
%\end{equation}

@@

\begin{equation}[!h]
\label{eq:CMatrix}
 C(z_0,z_n) =  \prod_{i=1}^{n}\me^{S_{1,i}(z_i-z_{i-1})}\me^{S_{2,i}(z_i-z_{i-1})}
\end{equation}

@@

The device was fabricated using a CMOS-compatible technology with electron-beam lithography. 
Fiber grating couplers were used as optical IOs in the measurement. 
Fig.~\ref{fig:passive} shows the measured drop port response of the cascaded contra-DC filters. 
The measurements were normalized using the response of a pair of directly connected fiber grating couplers on the same chip.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Passive6}
\caption{ Spectral response of the cascaded filters without heating: 1-dB BW is 733 GHz, 3-dB BW is 788 GHz, 10-dB BW is 868 GHz, and 20-dB BW is 990 GHz.}
\label{fig:passive}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Band5}
\caption{Spectral response of the device for different temperatures applied to only one contra-DC: 1-dB BW tuned down to 65 GHz, 3-dB BW to 117 GHz. The temperatures are obtained from simulation fit.}
\label{fig:bandTune}
\end{figure}

@@

%\section{Conclusion}
In summary, we have demonstrated a bandwidth tunable filter with a low insertion loss of less then 0.5 dB, low ripples of less than 0.3 dB, a large maximal bandwidth of greater than 750 GHz, and high contrast of 55 dB. 
Compared to other devices based on microring resonators and MZIs, it enables a much larger maximal passband and does not suffer from small FSRs. 
A large bandwidth tuning range over 670 GHz has been achieved, which, to the best of our knowledge, is the widest ever demonstrated on a silicon chip. 
This ultra-wide bandwidth tunability makes the device very attractive for next-generation ultrahigh baud-rate applications (e.g., high-capacity super-channel transmissions) and flexible optical networking.

@@

    \hline
   \end{tabular}
    \end{table*}

>>>>

A broadband filter design for a coarse WDM on the 220-nm SOI platform \cite{shi2013siliconCWDM} is adopted for individual contra-DCs.  
In each contra-DC, the widths of the waveguides are of 560 and 440 nm.
This high asymmetry ensures a negligible forward coupling. 
The gratings are formed by sidewall corrugations in strip waveguides with a pitch of 312 nm. 
The spacing between the waveguides varies between 65 and 135 nm with a Gaussian profile as shown in Fig.~\ref{fig:SEM} for efficient sidelobe suppression, for which details can be found in \cite{shi2013siliconCWDM}.
The spectral responses of each contra-DC is calculated using coupled mode theory, following the same procedure as in  \cite{shi2013siliconContraDC}, with three additions: apodization, temperature and noise simulation.

%The first part in predicting the contra-directional coupler spectrum is to compute the dielectric variation caused by the perturbation $\Delta\varepsilon(x,y)$ which is $\varepsilon_0/\pi*(n_{max}(x,y)^2-n_{min}(x,y)^2)$ where the $1/\pi$ factor is the first Fourier expansion coefficient in the case of a square shaped perturbation. 
%One also needs the mode distribution in each waveguide  $E_1(x,y)$ and $E_2(x,y)$ at the average point of the perturbation, which can be obtained using a 2D eigenmode solver such as Lumerical MODE. 
%The couplings $\kappa_{11}$, $\kappa_{12}$ and $\kappa_{22}$  can then be calculated from
%\begin{equation}\label{eq:coupling}
%\kappa_{nm}=\frac{\omega}{4}\iint \textbf{E}^*_n(x,y)\Delta\varepsilon_1(x,y) \textbf{E}_m(x,y) \diff x \diff y
%\end{equation}

@@

\begin{equation}
\label{eq:CMatrix}
 C(z_0,z_n) =  \prod_{i=1}^{n}\me^{S_{1,i}(z_i-z_{i-1})}\me^{S_{2,i}(z_i-z_{i-1})}
\end{equation}

@@

The device was fabricated using a CMOS-compatible technology with electron-beam lithography. 
Fiber grating couplers\cite{zhong2014focusingFGC} were used as optical IOs in the measurement. 
Fig.~\ref{fig:passive} shows the measured drop port response of the cascaded contra-DC filters. 
The measurements were normalized using the response of a pair of directly connected fiber grating couplers on the same chip.
\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Passive6}
\caption{ Spectral response of the cascaded filters without heating: 1-dB BW is 733 GHz, 3-dB BW is 788 GHz, 10-dB BW is 868 GHz, and 20-dB BW is 990 GHz.}
\label{fig:passive}
\end{figure}

@@

\begin{figure}[htbp]
\centering
\includegraphics[width=.99\columnwidth]{data/Band6}
\caption{Spectral response of the device for different temperatures applied to only one contra-DC: 1-dB BW tuned down to 65 GHz, 3-dB BW to 117 GHz. The temperatures are obtained from simulation fit.}
\label{fig:bandTune}
\end{figure}

@@

%\section{Conclusion}
In summary, we have demonstrated a bandwidth tunable filter with a low insertion loss of less then 0.5 dB, low ripples of less than 0.3 dB, a large maximal bandwidth of greater than 750 GHz, and high contrast of 55 dB. 
Compared to other devices based on microring resonators and MZIs (see Table \ref{table:comparison}), it enables a much larger maximal passband and does not suffer from small FSRs. 
A large bandwidth tuning range over 670 GHz has been achieved, which, to the best of our knowledge, is the widest ever demonstrated on a silicon chip. 
This ultra-wide bandwidth tunability makes the device very attractive for next-generation ultrahigh baud-rate applications (e.g., high-capacity super-channel transmissions) and flexible optical networking.

@@

    \hline
    \label{table:comparison}
   \end{tabular}
    \end{table*}
